---
title: "DATA MINING COURSE"
author: "TEYI KODJO JEROME SEDOWO"
date: "`r Sys.Date()`"
output: word_document
---

# Package

**Here, we'll use FactoMineR for the analysis and factoextra for ggplot2-based visualization**

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(FactoMineR)
library(factoextra)
library(readr)
library(corrplot)
```


# Import the data set

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data <- read_delim("C:/Users/Jérôme/Desktop/TEYI_KODJO_JEROME_SEDOWO/ETUDE/AIMS_SENEGAL_2024-2025/Review phase Courses/Block 4/Data Mining and Big data/Tutorial_1/ACP_eaux.txt", 
    delim = "\t", escape_double = FALSE, 
    col_types = cols(CA = col_number(), MG = col_number(), 
        `NA` = col_number(), K = col_number(), 
        SUL = col_number(), NO3 = col_number(), 
        HCO3 = col_number(), CL = col_number()), 
    trim_ws = TRUE)
#View(data)
```



```{r echo=TRUE}
data_numeric <- data[6:13]
head(data_numeric)
```

# Exploratory analysis

Here we print a simple statistics for continous variables

```{r echo=TRUE}
summary(data_numeric)
```

```{r echo=TRUE}
cor(data_numeric)
```



# Data Standardization

By default PCA() in **FactoMinR** standardizes the data automatically during the PCA. So we will not standardize the data manually before the PCA

```{r echo=TRUE}
resul_pca <- PCA(data_numeric, graph = FALSE)
print(resul_pca)
```


*This is many information found in many different lists and matrices.*


# Visualization and Interpretation

### Eigenvalues

```{r}
eig_value <- get_eigenvalue(resul_pca)
eig_value
```

### Visualisation and Interpretation

```{r}
fviz_eig(resul_pca, addlabels = TRUE, ylim = c(0,50))
```

From yhe plot above, we might want to stop at the third principal component. 68% of the information contained in the data are retained by the first Three principal components.

## Graph of variables 

```{r}
var <- get_pca_var(resul_pca)
var
```

### Coordinates of variables

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
var$coord
```

## Contribution of the variables

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
var$contrib
```
## Contribution of Variables to PC1

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_contrib(resul_pca, choice = "var", axes = 1)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_contrib(resul_pca, choice = "var", axes = 2)
```


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_contrib(resul_pca, choice = "var", axes = 3)
```




```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_var(resul_pca, col.var = "contrib",
             gradient.cols = c("#00AFBB","#E7B800","#FC4E07"),
             repel = TRUE)
```





### Correlation circle

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_var(resul_pca, col.var = "cos2",
             gradient.cols = c("#00AFBB","#E7B800","#FC4E07"),
             repel = TRUE)
```


## Quality of representation

```{r}
corrplot(var$cos2, is.corr = FALSE)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_cos2(resul_pca, choice = "var", axes = 1:2)
```


## Dimension description 

*Description of dimension 1*

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
res.desc <- dimdesc(resul_pca, axes = c(1,2), proba = 0.05)
res.desc$Dim.1
```
*Description of dimension 2*

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
res.desc <- dimdesc(resul_pca, axes = c(1,2), proba = 0.05)
res.desc$Dim.2
```
# Graph of individuals 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

ind <- get_pca_ind(resul_pca)
ind
```
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# ind$coord
# ind$cos2
# ind$contrib

```


# Quality of contribution


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_ind(resul_pca, col.ind = "cos2",
             gradient.cols = c("#00AFBB","#E7B800","#FC4E07"),
             repel = TRUE)
```


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

fviz_pca_ind(resul_pca, pointsize = "cos2",
             pointshape = 21, fill = "#E7B800",
             repel = TRUE)
```


